GSE158422: DNA methylation in Lung squamous cell carcinoma - gene methylation enrichment analysis

### 2023-10-12 figs/GSE158422_split.Rmd
moving on to sensitivity analysis
add mitch competitive test

# optimised for 128 GB sever with 32 threads
CORES=8

@@

@@

aares <- mclapply( seq(100,300,100) ,splitanalyze_aa,mc.cores=CORES)

@@

@@

@@

res <- do.call(rbind,lapply(list(laout,alout,aaout),colMeans))

rownames(res) <- c("la","al","aa")

@@

apply(laout,2,mean)
apply(alout,2,mean)
apply(aaout,2,mean)

@@

barplot(res$common,main="no. common",names.arg=c("LA","AL","AA"))
barplot(res$uncommon,main="no. uncommon",names.arg=c("LA","AL","AA"))
barplot(res$discordant,main="no. discordant",names.arg=c("LA","AL","AA"))

boxplot(list("LA"=laout$common,"AL"=alout$common,"AA"=aaout$common),
  main="no. common",col="white",ylim=c(0,1550))
beeswarm(list("LA"=laout$common,"AL"=alout$common,"AA"=aaout$common), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$uncommon,"AL"=alout$uncommon,"AA"=aaout$uncommon),
  main="no. uncommon",col="white",ylim=c(0,1550))
beeswarm(list("LA"=laout$uncommon,"AL"=alout$uncommon,"AA"=aaout$uncommon), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$discordant,"AL"=alout$discordant,"AA"=aaout$discordant),
  main="no. discordant",col="white",ylim=c(0,1550))
beeswarm(list("LA"=laout$discordant,"AL"=alout$discordant,"AA"=aaout$discordant), add=TRUE,cex=1.2,pch=19)

barplot(res$p_comm,main="proportion common",names.arg=c("LA","AL","AA"))
barplot(res$p_uncom,main="proportion uncommon",names.arg=c("LA","AL","AA"))
barplot(res$p_disc,main="proportion discordant",names.arg=c("LA","AL","AA"))

boxplot(list("LA"=laout$p_comm,"AL"=alout$p_comm,"AA"=aaout$p_comm),
  main="proportion common",col="white",ylim=c(0,1))
beeswarm(list("LA"=laout$p_comm,"AL"=alout$p_comm,"AA"=aaout$p_comm), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$p_uncom,"AL"=alout$p_uncom,"AA"=aaout$p_uncom),
  main="proportion uncommon",col="white",ylim=c(0,1))
beeswarm(list("LA"=laout$p_uncom,"AL"=alout$p_uncom,"AA"=aaout$p_uncom), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$p_disc,"AL"=alout$p_disc,"AA"=aaout$p_disc),
  main="proportion discordant",col="white",ylim=c(0,1))
beeswarm(list("LA"=laout$p_disc,"AL"=alout$p_disc,"AA"=aaout$p_disc), add=TRUE,cex=1.2,pch=19)

>>>>

# optimised for 128 GB sever with 32 threads
CORES=6

@@

## Agg-limma-mitch function

This approach uses the aggregated mvals, limma and instead of a 1-sample t-test it uses mitch
which is a competitive test and could give more interpretable results.

```{r,alm}

runmitch <- function(m,genesets,cores=1) {
  mres <- mitch_calc(m,genesets,cores=cores)
  mres <- mres$enrichment_result
  rownames(mres) <- mres$set
  mres$set=NULL
  return(mres)
}

@@

aares <- mclapply( seq(100,5000,100) ,splitanalyze_aa,mc.cores=CORES)

## ALM competitive test

```{r,almflow}

splitanalyze_alm <- function(seed){
  set.seed(seed)
  a <- sample(1:37,18)
  a <- a[order(a)]
  nona <- setdiff(1:37,a)
  set.seed(seed)
  b <- sample(nona,18)
  b <- b[order(b)]
  a <- c( (a*2)-1 ,a*2 )
  a <- a[order(a)]
  b <- c( (b*2)-1 ,b*2 )
  b <- b[order(b)]

  #al
  medf1 <- chragg(mval_a,myann,cores=floor(CORES/3))
  magg1 <- agglimma(medf1,design_a)
  m1 <- as.data.frame(magg1$t)
  rownames(m1) <- rownames(magg1)
  colnames(m1) <- "t"
  almres1 <- runmitch(m=m1,genesets=gs_symbols,cores=CORES)

  #al
  medf2 <- chragg(mval_b,myann,cores=floor(CORES/3))
  magg2 <- agglimma(medf2,design_b)
  m2 <- as.data.frame(magg2$t)
  rownames(m2) <- rownames(magg2)
  colnames(m2) <- "t"
  almres2 <- runmitch(m=m2,genesets=gs_symbols,cores=CORES)

  list("almres1"=almres1,"almres2"=almres2)

almres <- mclapply( seq(100,5000,100) ,splitanalyze_alm,mc.cores=CORES)

@@

almcompare <-function(almres){
  almres1 <- almres[[1]]
  almres2 <- almres[[2]]
  almsig1 <- subset(almres1,p.adjustANOVA<0.05)
  almsig2 <- subset(almres2,p.adjustANOVA<0.05)

  almsig1_up <- rownames(subset(almsig1,s.dist>0))
  almsig1_dn <- rownames(subset(almsig1,s.dist<0))
  almsig2_up <- rownames(subset(almsig2,s.dist>0))
  almsig2_dn <- rownames(subset(almsig2,s.dist<0))

  total <- length(unique(c(almsig1_up,almsig2_up,almsig1_dn,almsig2_dn)))
  common <- length(intersect(almsig1_up, almsig2_up)) + length(intersect(almsig1_dn, almsig2_dn))
  discordant <- length(intersect(almsig1_up, almsig2_dn)) + length(intersect(almsig1_dn, almsig2_up))
  uncommon = total - common - discordant
  result <- c("total"=total,"common"=common,"discordant"=discordant,"uncommon"=uncommon)

  almout <- data.frame(total,common,discordant,uncommon)
  almout$p_comm <-almout$common / almout$total
  almout$p_disc <-almout$discordant / almout$total
  almout$p_uncom <-almout$uncommon / almout$total
  almout
}

@@

almout <- do.call(rbind,lapply(almres,alcompare))

@@

almout

res <- do.call(rbind,lapply(list(laout,alout,aaout,almout),colMeans))

rownames(res) <- c("la","al","aa","alm")

@@

apply(laout,2,mean)
apply(alout,2,mean)
apply(aaout,2,mean)
apply(almout,2,mean)

@@

barplot(res$common,main="no. common",names.arg=c("LA","AL","AA","ALM"))
barplot(res$uncommon,main="no. uncommon",names.arg=c("LA","AL","AA","ALM"))
barplot(res$discordant,main="no. discordant",names.arg=c("LA","AL","AA","ALM"))

boxplot(list("LA"=laout$common,"AL"=alout$common,"AA"=aaout$common,"ALM"=almout$common),
  main="no. common",col="white",ylim=c(0,1550))
beeswarm(list("LA"=laout$common,"AL"=alout$common,"AA"=aaout$common,"ALM"=almout$common), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$uncommon,"AL"=alout$uncommon,"AA"=aaout$uncommon,"ALM"=almout$uncommon),
  main="no. uncommon",col="white",ylim=c(0,1550))
beeswarm(list("LA"=laout$uncommon,"AL"=alout$uncommon,"AA"=aaout$uncommon,"ALM"=almout$uncommon), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$discordant,"AL"=alout$discordant,"AA"=aaout$discordant,"ALM"=almout$discordant),
  main="no. discordant",col="white",ylim=c(0,1550))
beeswarm(list("LA"=laout$discordant,"AL"=alout$discordant,"AA"=aaout$discordant,"ALM"=almout$discordant), add=TRUE,cex=1.2,pch=19)

barplot(res$p_comm,main="proportion common",names.arg=c("LA","AL","AA","ALM"))
barplot(res$p_uncom,main="proportion uncommon",names.arg=c("LA","AL","AA","ALM"))
barplot(res$p_disc,main="proportion discordant",names.arg=c("LA","AL","AA","ALM"))

boxplot(list("LA"=laout$p_comm,"AL"=alout$p_comm,"AA"=aaout$p_comm,"ALM"=almout$p_comm),
  main="proportion common",col="white",ylim=c(0,1))
beeswarm(list("LA"=laout$p_comm,"AL"=alout$p_comm,"AA"=aaout$p_comm,"ALM"=almout$p_comm), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$p_uncom,"AL"=alout$p_uncom,"AA"=aaout$p_uncom,"ALM"=almout$p_uncom),
  main="proportion uncommon",col="white",ylim=c(0,1))
beeswarm(list("LA"=laout$p_uncom,"AL"=alout$p_uncom,"AA"=aaout$p_uncom,"ALM"=almout$p_uncom), add=TRUE,cex=1.2,pch=19)

boxplot(list("LA"=laout$p_disc,"AL"=alout$p_disc,"AA"=aaout$p_disc,"ALM"=almout$p_disc),
  main="proportion discordant",col="white",ylim=c(0,1))
beeswarm(list("LA"=laout$p_disc,"AL"=alout$p_disc,"AA"=aaout$p_disc,"ALM"=almout$p_disc), add=TRUE,cex=1.2,pch=19)
